---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# SCRI

<!-- badges: start -->
[![R-CMD-check](https://github.com/UMC-Utrecht-RWE/SCRI/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/UMC-Utrecht-RWE/SCRI/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of this package is to aid researchers in preparing analytic datasets for Self Controlled Risk Interval Analysis.

## Installation

You can install the development version of SCRI from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("UMC-Utrecht-RWE/SCRI")
```

## Example

The package implements a processing pipeline, which takes a study population and *metadata* which specifies analysis options, and outputs an analytic dataset. The pipeline is described in schematic form below, with purple boxes representing functions of the pacakge, blue boxes representing intermediate outputs, red and orange represents user input in terms of specifying analysis options and (raw) input data respectively

![image](https://github.com/UMC-Utrecht-RWE/SCRI/assets/138911044/8cbb1033-3e7b-4a05-9c0f-904e55f9fea3)

The package comes with sample input files for the `StudyPopulation` and `WindowsMetadata` files respectively.

### Step 1: Computing Windows

In this step, we use the `compute_windows()` function to calculate the start and end dates of different analysis windows. Let's start by grabbing the relevant windows from our example metadata file

```{r}
# library(SCRI)
data("StudyPopulation")
data("WindowsMetadata")
windowmet <- WindowsMetadata[grepl("post",WindowsMetadata$window_name),]
print(windowmet)
```


We are going to use this object in combination with our `StudyPopulation`, which contains one row per unit of analysis, with information on the `reference_date` and other dates which will be relevant in future steps.

```{r}
head(StudyPopulation)
```

The `compute_windows()` returns start and end dates for each of the windows under consideration, anchored to the relevant reference date. The output by default is wide format. The output object also inherits all columns from the input `studypopulation`

```{r}
sp_windows <- SCRI::compute_windows(studypopulation = StudyPopulation, windowmeta= windowmet, 
                id_column = "person_id")

sp_windows[1,c("person_id", "start_risk_post","end_risk_post","start_control_post","end_control_post")]
```

#### Step 2: Cleaning Windows

Now that we have the start and end dates of each window, we can proceed to the second step: Applying different rules which **alter** or **clean** those start and end dates based on other available information. Currently we have implemented **censoring date** functionality according to two rules: 

(1) If a censoring event occurs *during* a window, then the end date of that window is moved to the censoring date. All windows that start after that censoring date have start_ and end_ dates set to NA.

(2) If a censoring event occurs *between* windows, all future start and end dates are set to NA.

The function `clean_windows()` takes as input a wide-format dataset following the same format as that created by `compute_windows(..., output_format = "wide")`. In addition, the function takes the names of different date columns in the input dataset which are used for censoring. 

```{r}
# example rows with a single censoring date
sp_windows[c(1,17),c("start_risk_post","end_risk_post","start_control_post","end_control_post", "general_end_fup")]

# run the core function
sp_clean <- SCRI::clean_windows(sp_windows, censoring_dates = c("death_date", "general_end_fup"))

# output object has altered both example rows according to rules (1) and (2) above
sp_clean[c(1,17),c("start_risk_post","end_risk_post","start_control_post","end_control_post", "general_end_fup")]
```


