---
title: "Getting Started with the Self-Control-Risk-Interval (SCRI) Analysis Package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with the SCRI Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{css,  echo = F}
/*-- Specify div's for 'boxes', change color of TOC and center align titles: --*/
div.box1 {background-color: #f5f5f0; border-radius: 5px; padding: 30px; margin-right: 0px}
div.box2 {background-color: #d4e5d2; border-radius: 5px; padding: 30px; margin-right: 0px}
div.box3 {border-style: solid; border-color: #f5f5f0; border-width: medium; border-radius: 30px; padding: 5px; margin-right: 0px}
 
.list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {background-color: #76b82a; border-color: #76b82a}
h1 {text-align: center; color: #3c7b8a}
h2 {text-align: center; color: #76b82a}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(SCRI)
```

<div class = 'box3'>
# Introduction

The **Self-Control-Risk-Interval (SCRI)** R package provides a pipeline apply the Self-Controlled Risk Interval (SCRI) analysis in a modular and reproducible format. 

This vignette focuses on a vaccine safety setting aligned with the package manuscript introduction. The study question is whether the risk of acute events of special interest increases after COVID-19 vaccination compared with control periods within the same person.

For this worked example:
- Exposure anchors are `covid_vaccine_1` and `covid_vaccine_2`
- Outcomes of interest are `myocarditis` and `pericarditis`
- Censoring is applied using `death_date` and `general_end_fup`

# Installation

You can install the development version of the package from GitHub:

```{r eval=FALSE}
# install.packages("devtools")
library(devtools)
devtools::install_github("UMC-Utrecht-RWE/SCRI")
```

<br>

# Workflow Overview

The SCRI package workflow is as follows:

1.  **Prepare input data:** Study population, window metadata, and records tables.
2.  **Construct analysis windows:** Define risk, control, and other windows for each subject.
3.  **Apply censoring/cleaning rules:** Trim windows based on follow-up or censoring criteria.
4.  **Match events to windows:** Identify outcomes/events in each window.
5.  **Prepare the analysis dataset:** Summarize events and person-time.
6.  **Run analysis:** Fit conditional logistic regression or other models.
7.  **Summarize and interpret results.**

# Example Data

The package provides four example datasets needed to run an SCRI analysis:

**1. StudyPopulation:** Subject-level data with reference dates (e.g., prescription dates) and follow-up boundaries

```{r}
data("StudyPopulation")
DT::datatable(StudyPopulation, options = list(scrollX = TRUE))
```

**2. WindowMetadata:** Definitions of analysis windows (risk, control, washout, etc.) with their durations and timing relative to reference dates

```{r}
data("WindowMetadata")
DT::datatable(WindowMetadata, options = list(scrollX = TRUE))
```

**3. RecordsTable:** Outcome events (e.g., adverse events, hospitalizations) with dates for each subject

```{r}
data("RecordsTable")
DT::datatable(RecordsTable, options = list(scrollX = TRUE))
```

**4. WindowPriority:** Priority rules defining which window takes precedence when windows from consecutive doses overlap

```{r}
data("WindowPriority")
DT::datatable(WindowPriority, options = list(scrollX = TRUE))
```

# Step 1: Construct Windows

Calculate the start and end dates for each analysis window. The output is one row per dose/reference date, with a `t0_date` column that is used later to order consecutive doses.

**What happens:**
- For each person (id) and each reference date (covid_vaccine_1, covid_vaccine_2), windows are computed relative to that date
- `t0_date` is set equal to the reference date and used downstream to order doses chronologically
- Window definitions from `WindowMetadata` (e.g., risk, control, washout) are applied
- Each row contains start_ and end_ columns for every window type

```{r}
# Construct windows for all subjects relative to their covid_vaccine_1 and covid_vaccine_2 dates
constructed_windows <- construct_window(
    study_population = StudyPopulation, # Input: subject characteristics and reference dates
    windows_metadata = WindowMetadata, # Input: window definitions (duration, lag, etc.)
    reference_date_name = c("covid_vaccine_1", "covid_vaccine_2"), # Dates to anchor windows around
    id_column = "id" # Column identifying each subject
)

# Display first few rows with all computed window dates
DT::datatable(head(constructed_windows), options = list(scrollX = TRUE))
```

**Output structure:** Each row represents one subject-dose combination. If a subject has 2 reference dates (covid_vaccine_1 and covid_vaccine_2), there will be 2 rows per subject. Columns include `id`, `t0_date`, and start_/end_ pairs for each window (e.g., start_risk, end_risk, start_control, end_control).

# Step 2: Clean/Trim Windows

Apply censoring rules (e.g., end of follow-up or death) and overlap-priority rules between consecutive doses ordered by `t0_date`.

**What happens:**
1. **Censoring:** If a person dies or reaches end-of-follow-up, any window end dates after that censoring date are trimmed to the censoring date
2. **Overlap resolution:** When a person has multiple doses (e.g., two prescriptions), windows from dose N and dose N+1 may overlap. Priority rules define which dose's window takes precedence:
   - Priority "A": Dose N window prevails → dose N+1 window start is shifted forward
   - Priority "B": Dose N+1 window prevails → dose N window end is trimmed backward

The priority rules define all possible combinations between window types. Within each cell, "A" represents priority of dose_n over dose_n_plus_1 and "B" represents the opposite.

```{r}
# Display priority rules: which window takes precedence in overlaps
DT::datatable(WindowPriority, options = list(scrollX = TRUE))
```

```{r}
# Apply censoring and overlap-priority rules to clean the windows
wrangled_windows <- wrangle_window(
    constructed_windows, # Input: windows from Step 1
    censoring_dates = c("death_date", "general_end_fup"), # Trim windows at these dates
    windows_priority = WindowPriority # Apply priority rules for N vs N+1 overlaps
)

# Display cleaned windows (dates may be trimmed or shifted compared to constructed_windows)
DT::datatable(wrangled_windows, options = list(scrollX = TRUE))
```

**Output changes:** 
- Window end dates are clamped to censoring dates if applicable
- For subjects with multiple doses, overlapping windows are resolved per priority rules
- Window dates may now have NAs if they fall entirely after a censoring event

# Step 3: Match Records to Windows

Identify which outcomes/events occurred within each window.

**What happens:**
- Each record (outcome event) in `RecordsTable` is matched to the windows in `wrangled_windows`
- If a record's date falls within a window's start and end dates, it is flagged as occurring in that window
- The `add_records()` function creates indicators for each window showing whether the outcome occurred there
- Option: Keep only the first event per person, or retain all events

```{r}
# Match each outcome event to the appropriate window(s)
identified_records <- add_records(
    window_data = wrangled_windows, # Input: cleaned windows with adjusted dates
    records = RecordsTable, # Input: outcome events with dates
    only_first_record = TRUE # TRUE: keep only first event per subject; FALSE: keep all
)

# Display records with window assignments (new columns indicate which window each event occurred in)
DT::datatable(identified_records, options = list(scrollX = TRUE))
```

**Output structure:**
- Original window columns are retained
- New indicator columns show presence/absence of events in each window
- If stratification is needed, those columns are also present
- Rows without any matching records may be retained (useful for "unexposed" person-windows)

# Step 4: Prepare Analytical Dataset

Summarize events and person-time per window for analysis.

**What happens:**
- Event counts and person-time are aggregated by window, outcome type, and (optionally) stratification variables
- Outcome indicators are rolled up: if any event occurred in a window, that person-window is marked as an outcome
- Person-time is calculated as the number of days in each window (end_date - start_date)
- The dataset is restructured to match the format required by conditional logistic regression

```{r}
# Aggregate to create analytical dataset with events and person-time
analytical_dataset <- prepare_analytical_dataset(
    SCRI_identified_records = identified_records, # Input: records matched to windows
    strata_column_name = NULL, # Optional: column for stratification (or NULL for no stratification)
    only_first_date = TRUE # TRUE: use only first occurrence per person; FALSE: use all
)

# Display analytical dataset (one row per person-window combination)
DT::datatable(analytical_dataset, options = list(scrollX = TRUE))
```

**Output structure:**
- Rows represent person-window-outcome combinations
- Columns include window identifiers, outcome status (1/0), person-time, and any stratification variables
- This format is ready for conditional logistic regression
- Each subject appears once per window type (e.g., once for risk window, once for control window)

# Step 5: Run the Analysis

Apply the statistical model to estimate exposure effects.

**What happens:**
- A conditional logistic regression is fit with event status (outcome/no outcome) as the dependent variable
- The model conditions on each subject to account for person-level confounding (self-controlled design)
- The reference window (e.g., "control") defines unexposed person-time
- Risk and other windows define exposure periods
- The model compares odds of outcome during risk window vs. control window, within each person

```{r}
# Fit conditional logistic regression comparing windows within persons
results <- apply_analysis(
    SCRI_analytical_dataset = analytical_dataset, # Input: aggregated analytical dataset
    reference_date_name = "covid_vaccine_1", # Which reference date to use
    reference_window = "risk", # Which window serves as the baseline (usually "control")
    strata_column_name = NULL # Optional: column for stratified analysis
)

# Display results (odds ratios, confidence intervals, p-values)
DT::datatable(results, options = list(scrollX = TRUE))
```

**Output structure:**
- Columns include parameter estimates, odds ratios (ORs), confidence intervals (CIs), and p-values
- One row per window comparison (e.g., "risk" vs. reference window)
- ORs > 1 indicate increased risk during that window; ORs < 1 indicate decreased risk
- Statistical significance determined by whether CI excludes 1.0

# All-in-One Pipeline Function

You can use the main `scri()` function to run the entire pipeline end-to-end in a single call.

**What this does:**
- Internally performs all 5 steps above (construct → clean → match → prepare → analyze)
- Useful for quick analyses or when you don't need intermediate outputs
- Less flexible than stepwise approach, but more concise

```{r}
# Run complete SCRI analysis in one function call
final_results <- SCRI::scri(
    study_population = StudyPopulation, # Subject data with reference dates
    window_metadata = WindowMetadata, # Window definitions
    window_priority = WindowPriority, # Priority rules for dose overlaps
    records_table = RecordsTable, # Outcome events
    reference_date_name = "covid_vaccine_1", # Which reference date to use
    reference_window = "control", # Baseline window for comparison
    start_followup_criteria = "op_start_date", # Window construction start point
    end_followup_criteria = c("death_date", "general_end_fup"), # Censoring dates
    only_first_date = TRUE # Keep only first event per person
)

# Display final analysis results
DT::datatable(final_results, options = list(scrollX = TRUE))
```

**Advantages:**
- Single function call replaces 5 separate steps
- Automatically handles data flow between functions
- All internal steps use the same validation and error checking

**Disadvantages:**
- Cannot inspect intermediate outputs (windows before/after cleaning, records matching)
- Less flexibility for custom processing between steps

# Supported Study Designs and Models

You can query available self-controlled designs and valid models:

```{r}
get_ValidModels()
```

# Customization

The SCRI pipeline is highly customizable at multiple levels:

-   **Window definitions:** 
    - Edit `WindowMetadata` to change risk/control period durations, lags, and structure
    - Example: Extend washout period from 7 to 14 days by modifying the `duration` or `lag` columns
    
-   **Priority rules:** 
    - Modify `WindowPriority` to define how overlaps between consecutive doses are resolved
    - Example: Switch priority "A" ↔ "B" to give dose N+1 precedence instead of dose N
    
-   **Stratification:** 
    - Add stratification columns to `StudyPopulation` (e.g., age group, sex)
    - Specify the column name in `SCRI()` or `apply_analysis()` to run separate models per stratum
    
-   **Censoring criteria:** 
    - Use any date column in `StudyPopulation` (e.g., "death_date", "end_of_followup", "emigration_date")
    - Specify multiple censoring columns; windows are trimmed at the earliest censoring event
    
-   **Time-varying covariates:** 
    - Advanced: Use `apply_timevarying()` to include covariates that change over follow-up
    - Note: Experimental feature, consult documentation for details

-   **Event handling:**
    - Set `only_first_date = TRUE` to analyze only first outcome occurrence
    - Set `only_first_date = FALSE` to analyze all occurrences

# Conclusion

The SCRI package provides a comprehensive, modular, and reproducible workflow for conducting self-controlled risk interval analysis:

- **Flexible design:** Choose between stepwise control (5 individual steps) or all-in-one pipeline, depending on your needs
- **Customizable:** Adapt windows, priorities, censoring, and stratification to your study design
- **Transparent:** Each step produces interpretable outputs suitable for sensitivity analyses and validation
- **Self-controlled framework:** Automatically accounts for time-invariant confounding through matched analysis design

For more information, advanced usage, and troubleshooting, please see:
- Package documentation: `?SCRI` and individual function help pages
- GitHub repository: UMC-Utrecht-RWE/SCRI
- Published methods paper: [Reference as applicable]

</div>
