---
title: "Getting Started with the Self-Control-Risk-Interval (SCRI) Analysis Package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with the SCRI Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{css,  echo = F}
/*-- Specify div's for 'boxes', change color of TOC and center align titles: --*/
div.box1 {background-color: #f5f5f0; border-radius: 5px; padding: 30px; margin-right: 0px}
div.box2 {background-color: #d4e5d2; border-radius: 5px; padding: 30px; margin-right: 0px}
div.box3 {border-style: solid; border-color: #f5f5f0; border-width: medium; border-radius: 30px; padding: 5px; margin-right: 0px}
 
.list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {background-color: #76b82a; border-color: #76b82a}
h1 {text-align: center; color: #3c7b8a}
h2 {text-align: center; color: #76b82a}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(SCRI)
```

<div class = 'box3'>
# Introduction

The **Self-Control-Risk-Interval (SCRI)** R package provides a pipeline apply the Self-Controlled Risk Interval (SCRI) analysis in a modular and reproducible format. 

# Installation

You can install the development version of the package from GitHub:

```{r eval=FALSE}
# install.packages("devtools")
library(devtools)
devtools::install_github("UMC-Utrecht-RWE/SCRI")
```

<br>

# Workflow Overview

The SCRI package workflow is as follows:

1.  **Prepare input data:** Study population, window metadata, and records tables.
2.  **Construct analysis windows:** Define risk, control, and other windows for each subject.
3.  **Apply censoring/cleaning rules:** Trim windows based on follow-up or censoring criteria.
4.  **Match events to windows:** Identify outcomes/events in each window.
5.  **Prepare the analysis dataset:** Summarize events and person-time.
6.  **Run analysis:** Fit conditional logistic regression or other models.
7.  **Summarize and interpret results.**

# Example Data

The package provides example datasets:

```{r}
data("StudyPopulation")
data("WindowsMetadata")
data("RecordsTable")
DT::datatable(StudyPopulation, options = list(scrollX = TRUE))
DT::datatable(WindowsMetadata, options = list(scrollX = TRUE))
DT::datatable(RecordsTable, options = list(scrollX = TRUE))
```

# Step 1: Construct Windows

Calculate the start and end dates for each analysis window. The output is one row per dose/reference date, with a `t0_date` column that is used later to order consecutive doses:

```{r}
constructed_windows <- construct_windows(
    study_population = StudyPopulation,
    windows_metadata = WindowsMetadata,
    reference_date_name = c("FIRST_TARGET", "SECOND_TARGET"),
    id_column = "id"
)
DT::datatable(head(constructed_windows), options = list(scrollX = TRUE))
```

# Step 2: Clean/Trim Windows

Apply censoring rules (e.g., end of follow-up or death) and overlap-priority rules between consecutive doses ordered by `t0_date`.

The priority rules accross windows of subsequent doses are defined in a matrix with the following format (see table below)
In the priority matrix we have all possible combinations between tables. Within each cell we need to define the priority between the windows, A represent priority of dose_n over dose_n_plus_1 and B represents the opposite.

```{r}
priority_table <- readRDS("../data/WindowsPriority.rds")
DT::datatable(priority_table, options = list(scrollX = TRUE))
```

```{r}
cleaned_windows <- wrangle_window(
    constructed_windows,
    censoring_dates = c("death_date", "general_end_fup"),
    windows_priority = priority_table
)
DT::datatable(cleaned_windows, options = list(scrollX = TRUE))
```

# Step 3: Match Records to Windows

Identify which outcomes/events occurred within each window:

```{r}
identified_records <- add_records(
    window_data = cleaned_windows,
    records = RecordsTable,
    only_first_record = TRUE # or FALSE to keep all
)
DT::datatable(identified_records, options = list(scrollX = TRUE))
```

# Step 4: Prepare Analytical Dataset

Summarize events per person, window, etc.:

```{r}
analytical_dataset <- prepare_analytical_dataset(
    SCRI_identified_records = identified_records,
    strata_column_name = NULL, # or specify a column for stratification
    only_first_date = TRUE
)
DT::datatable(analytical_dataset, options = list(scrollX = TRUE))
```

# Step 5: Run the Analysis

Apply the statistical model (e.g., conditional logistic regression):

```{r}
results <- apply_analysis(
    SCRI_analytical_dataset = analytical_dataset,
    reference_window = "clean_lookback_pre_FIRST_TARGET",
    strata_column_name = NULL # or specify for stratified analysis
)
DT::datatable(results, options = list(scrollX = TRUE))
```

# All-in-One Pipeline Function

You can use the main `SCRI()` function to run the pipeline end-to-end:

```{r eval=FALSE}
final_results <- SCRI(
    study_population = StudyPopulation,
    windows_metadata = WindowsMetadata,
    records_table = RecordsTable,
    SCRI_analysis_name = "scri",
    reference_date_name = "FIRST_TARGET",
    reference_window = "clean_lookback_pre_FIRST_TARGET",
    start_followup_criteria = "op_start_date",
    end_followup_criteria = c("death_date", "general_end_fup"),
    only_first_date = TRUE
)
```

# Supported Study Designs and Models

You can query available self-controlled designs and valid models:

```{r}
get_valid_models()
```

# Customization

-   **Window definitions:** Edit `WindowsMetadata` to change risk/control periods.
-   **Stratification:** Add columns to `StudyPopulation` and specify in `SCRI()`/`apply_analysis()`.
-   **Time-varying covariates:** Advanced use via `apply_timevarying()` (experimental).

# Conclusion

The SCRI package offers a modular, reproducible workflow for self-controlled epidemiological studies. For more information, see the package documentation and GitHub repository.

\`\`\`
</div>
