---
title: "Getting Started with the Self-Control-Risk-Interval (SCRI) Analysis Package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with the SCRI Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{css,  echo = F}
/*-- Specify div's for 'boxes', change color of TOC and center align titles: --*/
div.box1 {background-color: #f5f5f0; border-radius: 5px; padding: 30px; margin-right: 0px}
div.box2 {background-color: #d4e5d2; border-radius: 5px; padding: 30px; margin-right: 0px}
div.box3 {border-style: solid; border-color: #f5f5f0; border-width: medium; border-radius: 30px; padding: 5px; margin-right: 0px}
 
.list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {background-color: #76b82a; border-color: #76b82a}
h1 {text-align: center; color: #3c7b8a}
h2 {text-align: center; color: #76b82a}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(SCRI)
```

<div class = 'box3'>
# Introduction

The **Self-Control-Risk-Interval (SCRI)** R package provides a pipeline for preparing and analyzing datasets using self-controlled study designs such as the Self-Controlled Risk Interval (SCRI), Self-Controlled Case Series (SCCS), and related methods. The package is especially useful for pharmacoepidemiology and vaccine safety studies, and can be adapted for other applications.

# Installation

You can install the development version of the package from GitHub:

```{r eval=FALSE}
# install.packages("devtools")
library(devtools)
devtools::install_github("UMC-Utrecht-RWE/Self-Control-Risk-Interval")
```

<br>

# Workflow Overview

The SCRI package workflow is as follows:

1.  **Prepare input data:** Study population, window metadata, and records tables.
2.  **Construct analysis windows:** Define risk, control, and other windows for each subject.
3.  **Apply censoring/cleaning rules:** Trim windows based on follow-up or censoring criteria.
4.  **Match events to windows:** Identify outcomes/events in each window.
5.  **Prepare the analysis dataset:** Summarize events and person-time.
6.  **Run analysis:** Fit conditional logistic regression or other models.
7.  **Summarize and interpret results.**

# Example Data

The package provides example datasets:

```{r}
data("StudyPopulation")
data("WindowsMetadata")
data("RecordsTable")
DT::datatable(StudyPopulation,options = list(scrollX = TRUE))
DT::datatable(WindowsMetadata,options = list(scrollX = TRUE))
DT::datatable(RecordsTable,options = list(scrollX = TRUE))
```

# Step 1: Construct Windows

Calculate the start and end dates for each analysis window:

```{r}
contructed_windows <- Construct_windows(
  study_population = StudyPopulation,
  windows_metadata = WindowsMetadata,
  reference_date_name = "FIRST_TARGET",
  id_column = "id"
)
DT::datatable(contructed_windows,options = list(scrollX = TRUE))
```

# Step 2: Clean/Trim Windows

Apply censoring rules (e.g., end of follow-up or death):

```{r}
cleaned_windows <- wrangle_window(
  contructed_windows,
  censoring_dates = c("death_date", "general_end_fup")
)
DT::datatable(cleaned_windows,options = list(scrollX = TRUE))
```

# Step 3: Match Records to Windows

Identify which outcomes/events occurred within each window:

```{r}
identified_records <- add_records(
  window_data = cleaned_windows,
  records = RecordsTable,
  only_first_record = TRUE # or FALSE to keep all
)
DT::datatable(identified_records,options = list(scrollX = TRUE))
```

# Step 4: Prepare Analytical Dataset

Summarize events per person, window, etc.:

```{r}
analytical_dataset <- prepare_analytical_dataset(
  SCRI_identified_records = identified_records,
  strata_column_name = NULL, # or specify a column for stratification
  only_first_date = TRUE
)
DT::datatable(analytical_dataset,options = list(scrollX = TRUE))
```

# Step 5: Run the Analysis

Apply the statistical model (e.g., conditional logistic regression):

```{r}
results <- apply_analysis(
  SCRI_analytical_dataset = analytical_dataset,
  reference_window = 'clean_lookback_pre_FIRST_TARGET',
  strata_column_name = NULL # or specify for stratified analysis
)
DT::datatable(results,options = list(scrollX = TRUE))
```

# All-in-One Pipeline Function

You can use the main `SCRI()` function to run the pipeline end-to-end:

```{r eval=FALSE}
final_results <- SCRI(
  study_population = StudyPopulation,
  windows_metadata = WindowsMetadata,
  records_table = RecordsTable,
  SCRI_analysis_name = "scri",
  reference_date_name = "FIRST_TARGET",
  reference_window = 'clean_lookback_pre_FIRST_TARGET',
  start_followup_criteria = "op_start_date",
  end_followup_criteria = c("death_date", "general_end_fup"),
  only_first_date = TRUE
)
```

# Supported Study Designs and Models

You can query available self-controlled designs and valid models:

```{r}
get_valid_SCRI()
get_valid_models()
```

# Customization

-   **Window definitions:** Edit `WindowsMetadata` to change risk/control periods.
-   **Stratification:** Add columns to `StudyPopulation` and specify in `SCRI()`/`apply_analysis()`.
-   **Time-varying covariates:** Advanced use via `apply_timevarying()` (experimental).

# Conclusion

The SCRI package offers a modular, reproducible workflow for self-controlled epidemiological studies. For more information, see the package documentation and GitHub repository.

\`\`\`
</div>
